An efficient acyclic contact planner for legged robots

### 2017-01-18 body/abstract.tex
andrea s comment + reviewers

% !TEX root =  ../main.tex
\adnote{What about replacing 'multiped' with 'legged' everywhere?}
\repadp{We present a contact planning framework for complex legged locomotion tasks:}{We present a framework capable of producing contact plans describing complex multiped motions}: standing up, climbing stairs using a handrail, crossing rubble and getting out of a car. The need for such a planner was highlighted at the Darpa Robotics Challenge, where \deladp{required} multi-contact behaviors
could not be demonstrated.

Indeed, current planners suffer from prohibitive computation times, \repadp{due to}{explained by} the difficulty
of generating contact postures: the feasible space (the contact manifold) is of high dimension,
and impossible to sample.

We tackle this issue with our novel \deladp{key} idea, the reachability condition. This condition allows us to define
a geometric approximation of the contact manifold, which is of low dimension and can be sampled efficiently.
Informally, the condition verifies that the root configuration of a robot is €œclose, but not too close to obstacles: close to allow contact creation, not too close to avoid collision. Thanks to this condition we decompose the hard contact planning problem into two
simpler sub-problems: first, planning a guide path for the robot root without considering the whole-body configuration, using a sampling-based algorithm; then, generating a discrete sequence of whole-body configurations in static equilibrium along this path, using a new deterministic
contact-selection algorithm.

Our approach results from the pragmatic choice of favoring efficiency over completeness, \repadp{which we justify empirically}{justified empirically}: in a
few seconds, with satisfying success rates, we generate complex contact plans for various scenarios and robots, namely
HRP-2, HyQ, and a dexterous hand. These plans are then validated either in dynamic simulations, or on the real HRP-2 robot, thus demonstrating 
the first interactive multi-contact planner.

>>>>

% !TEX root =  ../main.tex
We present a contact planning framework for complex legged locomotion tasks: standing up, climbing stairs using a handrail, crossing rubble and getting out of a car. The need for such a planner was highlighted at the Darpa Robotics Challenge, where multi-contact behaviors
could not be demonstrated.

Indeed, current planners suffer from prohibitive computation times, due to the difficulty
of generating contact postures: the feasible space (the contact manifold) is of high dimension,
and impossible to sample.

We tackle this issue with our novel  idea, the reachability condition. This condition allows us to define
a geometric approximation of the contact manifold, which is of low dimension and can be sampled efficiently.
Informally, the condition verifies that the root configuration of a robot is close, but not too close to obstacles: close to allow contact creation, not too close to avoid collision. Thanks to this condition we decompose the hard contact planning problem into two
simpler sub-problems: first, planning a guide path for the robot root without considering the whole-body configuration, using a sampling-based algorithm; then, generating a discrete sequence of whole-body configurations in static equilibrium along this path, using a new deterministic
contact-selection algorithm.

Our approach results from the pragmatic choice of favoring efficiency over completeness, which we justify empirically: in a
few seconds, with satisfying success rates, we generate complex contact plans for various scenarios and robots, namely
HRP-2, HyQ, and a dexterous hand. These plans are then validated either in dynamic simulations, or on the real HRP-2 robot, thus demonstrating 
the first interactive multi-contact planner.

### 2017-01-18 body/conclusion.tex
andrea s comment + reviewers

As demonstrated in the results section, the main interest of our method is time performance. Time performance is critical \deladp{to consider using a multi-contact planner} for 
online applications with the robot, a goal not satisfied by any of the previous contributions. Our method addresses highly constrained environments while improving the search time by orders of magnitude. This high performance is reached at the cost of some approximations that we discuss here. 
%~ In previous approaches, we identified 2 sources of high computational cost.

The first approximation is the verification of \contactreachability\ ($\mathbf{q}^0 \in C_{Contact}^0$).  Our \textit{reachability condition} ($\mathbf{q}^0 \in C_{Reach}^0$) is computationally efficient and provides an accurate approximation of $C_{Contact}^0$ (Section~\ref{sec:scaling}). This is demonstrated by the second column of Table~\ref{tab:requestpercent}, and illustrated by Figure~\ref{fig:dedefeas}. Indeed, in the large majority of cases, (84\% in the worst case \adnote{I do not get where this 84 comes from}), we are able to find a contact configuration for any configuration in $C_{Reach}^0$.

@@

\subsection{Addressing $\mathcal{P}_3$}
This paper focuses on \deladp{the} contact planning \deladp{aspect} of multi-contact locomotion. While it does not address directly the continuous motion generation,
we had to prove that the plans produced by our planner \deladp{are useful and} can be executed by the considered robots. Problem $\mathcal{P}_3$ remains \repadp{an active research topic~\cite{Carpentier2016, herzog2015trajectory}}{open},
and most existing solutions do not consider collisions with the environment during the planning, making it hard to validate constrained problems such as the polaris scenario
or the table egress in \cite{DBLP:conf/iser/EscandeKMG08}.
The companion video shows the results that we obtained in simulation and on the real HRP-2 robot. Despite the fact that our method does not consider joint torque limits in the planning
phase, the robot is able to execute our plans, achieving to demonstrate the interest of our framework.
\adnote{Probably you should add in appendix an explanation of how you solve $P_3$ and refer to it here.}

@@

%~ Our contribution to \Pa is a generic characterization of the properties that the guide path must satisfy, in particular to enforce the completeness of the acyclic contact planner. 
Our contribution to \Pb is a fast contact generation scheme that can \repadp{optimize user-defined criteria}{take into
account criteria to optimize user-defined properties}.

Our results demonstrate that our method allows a pragmatic compromise between three 
criteria that are hard to conciliate: generality, performance, and quality of the solution, making it the first acyclic contact
planner compatible with \gls{interactive} applications.
%
\textbf{Regarding generality}, the \textit{reachability condition}, coupled with an approach based on limb decomposition, 
allows the method to address automatically arbitrary multiped robots.
%
\textbf{Regarding performance}, our framework is \deladp{really} efficient in addressing both \Pa and $\mathcal{P}_2$. This results in \gls{interactive} computation times.
%
\textbf{Regarding the quality of the paths}, we are able to compute
\gls{equilibrium feasible} paths in all the presented scenarios, with high success rates.
As for \cite{Bouyarmane2009}, failures can still occur, due to the approximate condition used to compute the guide path.
The low computational burden of our framework however allows for fast re-planning in case of failure.
Furthermore, because of this approximation, the guide search is not complete. The choice is deliberate, because we \repadp{believe}{are convinced}
that it is necessary to trade completeness for efficiency at all stages of the planner.
However, one direction for future work is to focus on a more accurate formulation of $C_{reach}^0$ to improve
the approximation.

@@

While we have exhibited complete multi-contact locomotion obtained with our contact planner, our main concern for future work is to address the interpolation
between contact sequences ($\mathcal{P}_3$), not addressed here, and which remains an open issue in highly-constrained scenarios.
Solving $\mathcal{P}_3$ requires addressing efficiently the collision avoidance problem \deladp{during the interpolation},	but also improving
our plans with transition certificates, i.e. guarantees that the interpolation between two contact configurations is feasible.
%~ Our current direction is to propose a geometric quantification of both the kinematic and dynamic constraints of the problem, expressed
%~ at the center of mass of the robot.
 %~ We are also
%~ working on new heuristics more closely related to the dynamics of the systems (for instance, choose the contacts that minimize the jerk between two contact configurations).
Finally, we aim at performing kinodynamic planning to remove the constraint that configurations in the contact plan \deladp{should} be in static equilibrium. We believe that the most promising direction in this regard is to integrate the notion of Admissible Velocity Propagation in our current work \citep{DBLP:conf/rss/PhamCN13}.
Addressing these two issues is essential to bridge the gap between the planning and control aspects of multiped locomotion.

>>>>

As demonstrated in the results section, the main interest of our method is time performance. Time performance is critical for 
online applications with the robot, a goal not satisfied by any of the previous contributions. Our method addresses highly constrained environments while improving the search time by orders of magnitude. This high performance is reached at the cost of some approximations that we discuss here. 
%~ In previous approaches, we identified 2 sources of high computational cost.

The first approximation is the verification of \contactreachability\ ($\mathbf{q}^0 \in C_{Contact}^0$).  Our \textit{reachability condition} ($\mathbf{q}^0 \in C_{Reach}^0$) is computationally efficient and provides an accurate approximation of $C_{Contact}^0$ (Section~\ref{sec:scaling}). This is demonstrated by the second column of Table~\ref{tab:requestpercent}, and illustrated by Figure~\ref{fig:dedefeas}. Indeed, in the large majority of cases, (84\% in the worst car egress case), we are able to find a contact configuration for any configuration in $C_{Reach}^0$.

@@

\subsection{Addressing $\mathcal{P}_3$}
This paper focuses on contact planning of multi-contact locomotion. While it does not address directly the continuous motion generation,
we had to prove that the plans produced by our planner can be executed by the considered robots. Problem $\mathcal{P}_3$ remains an active research topic~\cite{Carpentier2016, herzog2015trajectory},
and most existing solutions do not consider collisions with the environment during the planning, making it hard to validate constrained problems such as the polaris scenario
or the table egress in \cite{DBLP:conf/iser/EscandeKMG08}.
The companion video shows the results that we obtained in simulation and on the real HRP-2 robot. Despite the fact that our method does not consider joint torque limits in the planning
phase, the robot is able to execute our plans, achieving to demonstrate the interest of our framework.
%~ \adnote{Probably you should add in appendix an explanation of how you solve $P_3$ and refer to it here.}

@@

%~ Our contribution to \Pa is a generic characterization of the properties that the guide path must satisfy, in particular to enforce the completeness of the acyclic contact planner. 
Our contribution to \Pb is a fast contact generation scheme that canoptimize user-defined criteria.

Our results demonstrate that our method allows a pragmatic compromise between three 
criteria that are hard to conciliate: generality, performance, and quality of the solution, making it the first acyclic contact
planner compatible with \gls{interactive} applications.
%
\textbf{Regarding generality}, the \textit{reachability condition}, coupled with an approach based on limb decomposition, 
allows the method to address automatically arbitrary multiped robots.
%
\textbf{Regarding performance}, our framework is efficient in addressing both \Pa and $\mathcal{P}_2$. This results in \gls{interactive} computation times.
%
\textbf{Regarding the quality of the paths}, we are able to compute
\gls{equilibrium feasible} paths in all the presented scenarios, with high success rates.
As for \cite{Bouyarmane2009}, failures can still occur, due to the approximate condition used to compute the guide path.
The low computational burden of our framework however allows for fast re-planning in case of failure.
Furthermore, because of this approximation, the guide search is not complete. The choice is deliberate, because we believe
that it is necessary to trade completeness for efficiency at all stages of the planner.
However, one direction for future work is to focus on a more accurate formulation of $C_{reach}^0$ to improve
the approximation.

@@

While we have exhibited complete multi-contact locomotion obtained with our contact planner, our main concern for future work is to address the interpolation
between contact sequences ($\mathcal{P}_3$), not addressed here, and which remains an open issue in highly-constrained scenarios.
Solving $\mathcal{P}_3$ requires addressing efficiently the collision avoidance problem,	but also improving
our plans with transition certificates, i.e. guarantees that the interpolation between two contact configurations is feasible.
%~ Our current direction is to propose a geometric quantification of both the kinematic and dynamic constraints of the problem, expressed
%~ at the center of mass of the robot.
 %~ We are also
%~ working on new heuristics more closely related to the dynamics of the systems (for instance, choose the contacts that minimize the jerk between two contact configurations).
Finally, we aim at performing kinodynamic planning to remove the constraint that configurations in the contact plan be in static equilibrium. We believe that the most promising direction in this regard is to integrate the notion of Admissible Velocity Propagation in our current work \citep{DBLP:conf/rss/PhamCN13}.
Addressing these two issues is essential to bridge the gap between the planning and control aspects of multiped locomotion.

### 2017-01-18 body/contact.tex
andrea s comment + reviewers

@@

%~ Figure~\ref{fig:break_contact} illustrates the contact-persistence strategy.
If kinematically possible, a limb in contact at step $i-1$ remains in contact at step $i$ (Figure~\ref{fig:break_contact}). 
%~ The contact is broken if an inverse-kinematics solver fails to find a collision-free limb configuration that satisfies joint limits. 
%~ The solver is directly provided by the HPP software.
Otherwise the contact is broken and a collision-free configuration is assigned to the limb.
If two or more contacts can't be maintained between two consecutive configurations, one or more intermediate configurations are added.
%~ For these steps the root configuration is the same as for the previous step, with the difference that
%~ one faulty contact is repositioned, in the hope that it will not be broken at the next step.

>>>>

\subsection{Definition of a contact sequence}
In previous contributions, a contact plan is defined as a sequence of configurations that describe a quasi-static equilibrium configuration
for each existing phase of the motion. For instance, a walk cycle would be described by three key configurations: a double support configuration,
a simple configuration (the contact is broken), and another double support configuration (a contact is created). Our definition of the contact plan differs: in this example it would simply consists in the two double support configurations. 
However, in the case where the contact broken and the contact created sequentially are not related to the same effector, our planner outputs two distinct states.
This representation is sufficient to describe all the contact constraints of the problem. Furthermore it removes the unnecessary constraint of having a simple support,
quasi static configuration in the path as in the example. As shown in the validation video, this allows our framework to produce motions that are not quasi static, but dynamic.

@@

%~ Figure~\ref{fig:break_contact} illustrates the contact-persistence strategy.
If kinematically possible, a limb in contact at step $i-1$ remains in contact at step $i$ (Figure~\ref{fig:break_contact}). 
%~ The contact is broken if an inverse-kinematics solver fails to find a collision-free limb configuration that satisfies joint limits. 
%~ The solver is directly provided by the HPP software.
Otherwise the contact is broken and a collision-free configuration is assigned to the limb.
If two or more contacts can't be maintained between two consecutive configurations, one or more intermediate configurations are added, to ensure
that at most one contact is broken between two sequential configurations.
%~ For these steps the root configuration is the same as for the previous step, with the difference that
%~ one faulty contact is repositioned, in the hope that it will not be broken at the next step.

### 2017-01-18 body/results.tex
andrea s comment + reviewers

% !TEX root =  ../main.tex
\section{Results}
\label{sec:results}
In this \repadp{section}{Section} we present some of the results obtained with our planner. The complete sequences \deladp{computed} are shown in the companion video.
Specifically, we demonstrate the planner for two really different robots, in a large variety of environments: the humanoid HRP-2 and the quadruped HyQ.
Finally, a last example suggests possible applications to dexterous manipulation.

@@

At the end of this \repadp{section}{Section}, we discuss the role of the parameters of our framework\deladp{ on the output}. We then provide the \textit{interactive} computation times obtained in each case.
We also compare the times obtained with HRP-2 with respect to previous works.

These scenarios \repadp{complement}{complete} the ones demonstrated on virtual avatars (Figure~\ref{fig:robots_old}) in our previous ISRR paper~\citep{tonneauisrr15} and video\footnote{\url{http://youtu.be/LmLAHgGQJGA}}.
%~ (Figure~\ref{fig:robots_old}).
 %~ We invite the interested reader to watch the ISRR video (\url{http://youtu.be/LmLAHgGQJGA}), and 
%~ to refer to the previous paper for a discussion on these results.

@@

\subsection{Description of the scenarios}
In all the scenarios considered, the formulation of the problem is always the same:
a start and goal root configurations are provided as \deladp{an} input \deladp{of the scenario}.
The framework computes the initial contact configuration, and outputs a sequence of contact configurations connecting it to the goal.
In each scenario we detail the contacts involved and the heuristics chosen (either $h_{\textrm{\it EFORT}}$, $h_{vel}$ or $h_{w}$, all of which are defined in the Appendix~\ref{sec:heuristics}).
 %~ and the constraints on the reachable workspaces (for instance in all the scenarios, the reachable workspaces of the legs of HRP-2 are always required to intersect with the environment). 
%~ The companion video presents the complete contact sequence obtained in all these scenarios.

@@

The goal is to climb three 15-cm high steps. This height requires HRP-2 to use a ramp to perform the task. \adnote{It is not clear why, so either explain why, or remove the comment.}

@@

\subsubsection{HRP-2 -- Standing up (Figure~\ref{fig:standing}):}
From a bent configuration, \repadp{the robot has to stand up}{a standing-up motion is computed in a constraining environment. The resulting motion involves} using a wall as support, and climbing a 25-cm high step.

@@

\subsubsection{HRP-2 -- Polaris car egress (Figure~\ref{fig:car}):}
\repadp{In this scenario---inspired from the DRC car egress (\url{http://cpc.cx/edH})---HRP-2 has}{This scenario is inspired from the DARPA-Robotics-Challenge car egress scenario (\url{http://cpc.cx/edH}). HRP-2 has to find the contact sequence that allows it} to step out of the Polaris car.

@@

\subsubsection{HyQ -- DRC-style rubble (Figure~\ref{fig:darpa})}
The quadruped robot \repadp{must}{is given the task to} cross a rubble composed of bricks rotated at different angles and directions.

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{figures/darpa}
  \caption{
           Robust crossing of rubbles by HyQ \deladp{($b_0 > 20$)}. }
		   \label{fig:darpa}
\end{figure}

@@

\subsubsection{HyQ -- Obstacle race (Figure~\ref{fig:HyQ_bridge} and~\ref{fig:HyQ_obs}):}
In this long scene, HyQ \repadp{has to cross a 55-cm large hole, followed by}{is first required to cross a 55-cm large hole; then, to cross} a narrow ``bridge'', only 25-cm large.

@@

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{figures/HyQ_obs}
  \caption{
           Crossing a hole contact sequence for HyQ \deladp{($b_0 > 4$)}. }
		   \label{fig:HyQ_obs}
\end{figure}

@@

\subsection{Role of the main parameters} \label{sec:influence}
We discuss the several factors \adnote{Either use 'the' or use 'several'} that influence the outcome of our planner: the root scaling factor $s$ (Section~\ref{sec:scaling}), the \deladp{choice of} heuristics for
contact generation (Appendix~\ref{sec:heuristics}), and lastly, the discretization step for the guide path. The appropriate value for these parameters
is computed empirically based on statistical analysis \adnote{Are you sure you wanna use the word 'statistical'?} or trials and errors.

\subsubsection{Choosing the scaling factor $s$:} \label{sec:params}
\deladp{To find a convenient value $s^*$ of $s$, we proceeded as follows. }
For several values of $s$, we generated $10 000$ configurations. 
We then computed the sensitivity of the reachability condition (\repadp{percentage}{proportion} of configurations in $C_{Reach}^0$, effectively belonging to \gls{$C_{Contact}^0$}).
Similarly we computed the specificity of the condition (\repadp{percentage}{proportion} of configurations \textbf{not} in $C_{Reach}^0$, effectively \textbf{not} belonging to \gls{$C_{Contact}^0$}).
The obtained results for HRP-2 are shown in Table~\ref{tab:scale}, averaged over all scenes (except for the car egress: in this scenario, 
statistical tests are not really conclusive since we are only interested in a small area of the environment).

As it can be expected, the scaling results in a high \repadp{increase}{decrease} of the sensitivity, with a decrease of the specificity.
For HRP-2 we decided to set $s^*=1.2$.
%~ , and was also effective for the car egress scenario.

\begin{table}
\centering
\footnotesize
\begin{tabular}{c | c | c}
   Value of $s$ &  Sensitivity & Specificity\\
 \hline
   1   & 76\% & 100 \%\\
   1.1& 88\% & 96\% \\
   1.15& 93\% & 94\%\\
   1.2 & 97\% & 92.5\%\\
   1.25& 98\% & 91.7\%\\
   1.5 & 99\% & 90.5\%\\
 \end{tabular}
\caption{Sensitivity and specificity values \addadp{of} the reachability condition, depending on the scaling value \addadp{$s$} of $W^0$.}
\label{tab:scale}
\quad
\end{table}

\subsubsection{\repadp{Choosing the heuristics}{On the choice of heuristics}:} \label{sec:heuristichoices}
In our conference paper~\citep{tonneauisrr15}, the computed motions were \deladp{only} generated using the EFORT heuristic.
EFORT is designed for tasks requiring to exert important forces (such as pushing / pulling / climbing). 
\repadp{In}{Regarding} locomotion tasks, such as the stair scenario, one issue with EFORT is that it tends to generate
configurations close to singularities (and joint limits). While this does not significantly impact
the generation of the plan, the resulting interpolation \deladp{on the real robot} turned out to be harder.
For this reason, we prefer to use our manipulability-based heuristic for the legs of the robot, but we still
use EFORT for the arms, which results in less contact repositionings.

@@

\subsubsection{Discretization of the guide path:} \label{sec:disc}
The discretization step is a user-defined, fixed parameter. The step
has an influence on the output of the planner: if too large steps are taken,
the planner may fail since we impose the constraint that only one contact change might occur
between two consecutive steps. On the other hand, a small step will not impact the success rate of the planner, 
but may generate unnecessary states. In most scenarios the torso of HRP-2 moves about 15 cm between two postures, but only 3 cm
for the car egress scenario to handle the geometry of the polaris.
For future work, we would like to automatically adapt the size of the discretization step to the complexity of the environment.
\adnote{Maybe you should say which discretization step you used.}

\subsection{Performance analysis} \label{sec:perf}
To analyze performance, we ran the planner 1000 times for each considered scenario.
We measured the computation time spent in each \repadp{part}{aspect} of the algorithm, and analyzed their success rate.

@@

\begin{table*}
\centering
\begin{tabular}{ l | c}
  &  Path planning success rate \\
 \hline
   Steep stairs     	& 100\% \\
   Standing up 			& 68\% 	\\
   Car egress 			& 77\% 	\\
   Rubble 				& 97\% 	\\
   Obstacle race        & 88.0\% 	\\
 \end{tabular}
\caption{ Percentage of successful complete contact planning rates for each scenario, rounded to the first decimal.}
\label{tab:sucess_planning}
\quad
\end{table*}

@@

\subsubsection{Computation times:}
Table~\ref{tab:requestime} summarizes the \deladp{performance measurements obtained, in terms of} computation times.

\repadp{F}{Considering the repartition of the computation time, f}or HRP-2, most of the time \repadp{was}{is} spent performing inverse kinematics.
This is not surprising considering the number of calls to the methods: IK projection is used intensively to maintain contact continuity between two postures; 
it is also applied every time a new candidate needs to be evaluated. In particular for the car egress scenario,
the kinematic constraints are very demanding to avoid collisions.

@@

In all scenarios, one can observe that the average computation time for one single step is largely below one second,
thus allowing to consider \gls{interactive} applications\repadp{ and}{. This would allow to consider} online autonomous planning of the robot motion.

\subsubsection{Success rates:}
Table~\ref{tab:sucess_planning} summarizes the successful planning rates\deladp{ obtained for each scenario}.
Despite the complexity of the scenarios addressed, as well as the approximations made in our formulation, our planner \repadp{succeeded}{succeeds} in the large majority of cases.

Table~\ref{tab:requestpercent} presents\deladp{, for each scenario,} the rate of \addadp{successful} contact generation\deladp{ that succeeded}. Note that a failure in contact generation for a root configuration is not equivalent to a failure in the contact plan. It simply means that another limb was tested for contact generation for the same root configuration.
As expected, the more constrained scenario, the car egress \deladp{one}, provides the \repadp{worst}{less satisfying} results, despite the high success rate of the planner.

These results confirm that our approach provides a satisfying compromise between completeness and efficiency, thus allowing to consider online planning
while controlling the robot. Indeed, when the contact planning fails, it fails rapidly. This allows us to rapidly re-plan with a reasonable chance of success.
The most efficient (and immediate) approach to obtain a valid contact plan as fast as possible \repadp{would be}{is probably} to launch in parallel several instances of the planner (our current implementation is single-threaded) and to use any successful result as a plan for solver $\mathcal{P}_3$.

>>>>

% !TEX root =  ../main.tex
\section{Results}
\label{sec:results}
In this section we present some of the results obtained with our planner. The complete sequences are shown in the companion video.
Specifically, we demonstrate the planner for two really different robots, in a large variety of environments: the humanoid HRP-2 and the quadruped HyQ.
Finally, a last example suggests possible applications to dexterous manipulation.

@@

At the end of this section, we discuss the role of the parameters of our framework. We then provide the \textit{interactive} computation times obtained in each case.
We also compare the times obtained with HRP-2 with respect to previous works.

These scenarios complement the ones demonstrated on virtual avatars (Figure~\ref{fig:robots_old}) in our previous ISRR paper~\citep{tonneauisrr15} and video\footnote{\url{http://youtu.be/LmLAHgGQJGA}}.
%~ (Figure~\ref{fig:robots_old}).
 %~ We invite the interested reader to watch the ISRR video (\url{http://youtu.be/LmLAHgGQJGA}), and 
%~ to refer to the previous paper for a discussion on these results.

@@

\subsection{Description of the scenarios}
In all the scenarios considered, the formulation of the problem is always the same:
a start and goal root configurations are provided as input.
The framework computes the initial contact configuration, and outputs a sequence of contact configurations connecting it to the goal.
In each scenario we detail the contacts involved and the heuristics chosen (either $h_{\textrm{\it EFORT}}$, $h_{vel}$ or $h_{w}$, all of which are defined in the Appendix~\ref{sec:heuristics}).
 %~ and the constraints on the reachable workspaces (for instance in all the scenarios, the reachable workspaces of the legs of HRP-2 are always required to intersect with the environment). 
%~ The companion video presents the complete contact sequence obtained in all these scenarios.

@@

The goal is to climb three 15-cm high steps.

@@

\subsubsection{HRP-2 -- Standing up (Figure~\ref{fig:standing}):}
From a bent configuration, the robot has to stand up using a wall as support, and climbing a 25-cm high step.

@@

\subsubsection{HRP-2 -- Polaris car egress (Figure~\ref{fig:car}):}
In this scenario inspired from the DRC car egress HRP-2 has to step out of the Polaris car.

@@

\subsubsection{HyQ -- DRC-style rubble (Figure~\ref{fig:darpa})}
The quadruped robot must cross a rubble composed of bricks rotated at different angles and directions.

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{figures/darpa}
  \caption{
           Robust crossing of rubbles by HyQ. }
		   \label{fig:darpa}
\end{figure}

@@

\subsubsection{HyQ -- Obstacle race (Figure~\ref{fig:HyQ_bridge} and~\ref{fig:HyQ_obs}):}
In this long scene, HyQ has to cross a 55-cm large hole, followed by a narrow ``bridge'', only 25-cm large.

@@

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{figures/HyQ_obs}
  \caption{
           Crossing a hole contact sequence for HyQ. }
		   \label{fig:HyQ_obs}
\end{figure}

@@

\subsection{Role of the main parameters} \label{sec:influence}
We discuss the factors that influence the outcome of our planner: the root scaling factor $s$ (Section~\ref{sec:scaling}), the heuristics for
contact generation (Appendix~\ref{sec:heuristics}), and lastly, the discretization step for the guide path. The appropriate value for these parameters
is computed empirically based on use-case analysis or trials and errors.

\subsubsection{Choosing the scaling factor $s$:} \label{sec:params}
%~ \deladp{To find a convenient value $s^*$ of $s$, we proceeded as follows. }
For several values of $s$, we generated $10 000$ configurations. 
We then computed the sensitivity of the reachability condition (percentage of configurations in $C_{Reach}^0$, effectively belonging to \gls{$C_{Contact}^0$}).
Similarly we computed the specificity of the condition (percentage of configurations \textbf{not} in $C_{Reach}^0$, effectively \textbf{not} belonging to \gls{$C_{Contact}^0$}).
The obtained results for HRP-2 are shown in Table~\ref{tab:scale}, averaged over all scenes (except for the car egress: in this scenario, 
statistical tests are not really conclusive since we are only interested in a small area of the environment).

As it can be expected, the scaling results in a high increasepercentage of the sensitivity, with a decrease of the specificity.
For HRP-2 we decided to set $s^*=1.2$.
%~ , and was also effective for the car egress scenario.

\begin{table}
\centering
\footnotesize
\begin{tabular}{c | c | c}
   Value of $s$ &  Sensitivity & Specificity\\
 \hline
   1   & 76\% & 100 \%\\
   1.1& 88\% & 96\% \\
   1.15& 93\% & 94\%\\
   1.2 & 97\% & 92.5\%\\
   1.25& 98\% & 91.7\%\\
   1.5 & 99\% & 90.5\%\\
 \end{tabular}
\caption{Sensitivity and specificity values of the reachability condition, depending on the scaling value $s$ of $W^0$.}
\label{tab:scale}
\quad
\end{table}

\subsubsection{Choosing the heuristics:} \label{sec:heuristichoices}
In our conference paper~\citep{tonneauisrr15}, the computed motions were generated using the EFORT heuristic.
EFORT is designed for tasks requiring to exert important forces (such as pushing / pulling / climbing). 
In locomotion tasks, such as the stair scenario, one issue with EFORT is that it tends to generate
configurations close to singularities (and joint limits). While this does not significantly impact
the generation of the plan, the resulting interpolation turned out to be harder.
For this reason, we prefer to use our manipulability-based heuristic for the legs of the robot, but we still
use EFORT for the arms, which results in less contact repositionings.

@@

\subsubsection{Discretization of the guide path:} \label{sec:disc}
The discretization step is a user-defined, fixed parameter. The step
has an influence on the output of the planner: if too large steps are taken,
the planner may fail since we impose the constraint that only one contact change might occur
between two consecutive steps. On the other hand, a small step will not impact the success rate of the planner, 
but may generate unnecessary states. In most scenarios the torso of HRP-2 moves about 15 cm between two postures, but only 3 cm
for the car egress scenario to handle the geometry of the polaris.
For future work, we would like to automatically adapt the size of the discretization step to the complexity of the environment.

\subsection{Performance analysis} \label{sec:perf}
To analyze performance, we ran the planner 1000 times for each considered scenario.
We measured the computation time spent in each part of the algorithm, and analyzed their success rate.

@@

\begin{table*}
\centering
\begin{tabular}{ l | c}
  &  Path planning success rate \\
 \hline
   Stairs     	& 100\% \\
   Standing			& 68\% 	\\
   Car 			& 77\% 	\\
   Rubble 				& 97\% 	\\
   Race        & 88.0\% 	\\
 \end{tabular}
\caption{ Percentage of successful complete contact planning rates for each scenario, rounded to the first decimal.}
\label{tab:sucess_planning}
\quad
\end{table*}

@@

\subsubsection{Computation times:}
Table~\ref{tab:requestime} summarizes the computation times.

For HRP-2, most of the time was spent performing inverse kinematics.
This is not surprising considering the number of calls to the methods: IK projection is used intensively to maintain contact continuity between two postures; 
it is also applied every time a new candidate needs to be evaluated. In particular for the car egress scenario,
the kinematic constraints are very demanding to avoid collisions.

@@

In all scenarios, one can observe that the average computation time for one single step is largely below one second,
thus allowing to consider \gls{interactive} applications and online autonomous planning of the robot motion.

\subsubsection{Success rates:}
Table~\ref{tab:sucess_planning} summarizes the successful planning rates.
Despite the complexity of the scenarios addressed, as well as the approximations made in our formulation, our planner succeeded in the large majority of cases.

Table~\ref{tab:requestpercent} presents the rate of successful contact generation. Note that a failure in contact generation for a root configuration is not equivalent to a failure in the contact plan. It simply means that another limb was tested for contact generation for the same root configuration.
As expected, the more constrained scenario, the car egress, provides the less satisfying results, despite the high success rate of the planner.

These results confirm that our approach provides a satisfying compromise between completeness and efficiency, thus allowing to consider online planning
while controlling the robot. Indeed, when the contact planning fails, it fails rapidly. This allows us to rapidly re-plan with a reasonable chance of success.
The most efficient (and immediate) approach to obtain a valid contact plan as fast as possible would be to launch in parallel several instances of the planner (our current implementation is single-threaded) and to use any successful result as a plan for solver $\mathcal{P}_3$.
